---
title: "Statistical Inference Project 1"
author: "Eric VACHON"
date: "May 2015"
output: pdf_document
geometry: margin = 1.2cm
---
```{r, echo=FALSE}
lambda <- 0.2
n <- 40
nbSimul <- 1000
```
## Overview
In this paper we explain and compare the distribution of `r nbSimul`  simulations of the mean of `r n` exponential distributions and the Central Limit Theorm (CLT). 
First we compare Sample Mean Versus Theorical Mean, then the Sample Variance 
Versus Theorical Variance and at end the difference between our simulation and the CLT.

## Requirement and reproductibility
Library : ggplot2 (If need install it : install.packages("ggplot2")) and for reproducability, we fixe the seed :
```{r,results='hide', message=FALSE, warning=FALSE}
library(ggplot2)
set.seed(19)  
```  
## Constants of the simulation
```{r}
lambda <- 0.2
n <- 40
nbSimul <- 1000
```
Results : 
- ${\lambda}$ =`r lambda` 
- n = `r n` 
- nbSimul= `r nbSimul` 
## Theoricals values 
Mean of exponencial distribution = $\frac{1}{\lambda}$  
Standard deviation of the mean of n exponencial distribution = $\frac{1}{\lambda}*\frac{1}{\sqrt{n}}$   
The variance of the mean of n exponencial distribution = $(\frac{1}{\lambda}*\frac{1}{\sqrt{n}})^2$
```{r}
theoricalMean <- (1 / lambda)
theoricalStandardDeviation <- (1 / lambda)*(1/sqrt(n))
theoricalVariance <- ((1 / lambda)*(1/sqrt(n)))^2
```
Results :   
- theoricalMean = `r theoricalMean`  
- theoricalStandardDeviation = `r theoricalStandardDeviation`  
- theoricalVariance = `r theoricalVariance`  

## The data frame of the simulation
Here we are going to make a data frames with the result of nbSimul=`r nbSimul` means of n=`r n` exponencial distribution

```{r}
mySimul <- data.frame()
for  (i in 1 : nbSimul) 
{  mySimul <-  rbind(mySimul,c(i,mean(rexp(n, lambda))))}
names(mySimul) <- c('numSimul','valueSimul')
```

## 1. Sample Mean Versus Theorical Mean
Now we make the histogram of this `r nbSimul` simulations and put a line of the 
theorical mean
```{r,results='hide', message=FALSE, warning=FALSE}
ggplot(mySimul, aes(x=valueSimul))+
   geom_histogram(aes(y=..density..), colour="black", fill="white")+
   geom_density(alpha=.2, fill="#000066")+
   geom_vline(xintercept=theoricalMean,color="red",linetype="dashed",size=2)+
   ggtitle("Sample mean Vs Theorical mean")+labs(x = "mean")
```  
  
And we can compare sample mean and theorical mean :
```{r}
sampleMean <- mean(mySimul$valueSimul)
```
Results :   
- sampleMean = `r sampleMean`   
- theoricalMean = `r theoricalMean`  
$\Rightarrow$ `r abs(round(100*(1-(sampleMean/theoricalMean)),2))`% of difference, so it is a good estimator.  

## 2. Sample Variance Versus Theorical Variance
Here we must compare the therical variuance and the variance of our simulation :  

```{r}
simulationVariance <- var(mySimul$valueSimul)
```
Results :   
- simulationVariance = `r simulationVariance`  
- theoricalVariance = `r theoricalVariance`  

$\Rightarrow$ `r abs(round(100*(1-simulationVariance/theoricalVariance),2))`% of difference, so it is a good estimator.

## 3. Comparaison with CLT  
Now let's use the formula of the CLT : $\frac{\sqrt n (\bar X_n - \mu)}{\sigma}$ and use n=10 and n=100 :  
```{r,results='hide', message=FALSE, warning=FALSE}
set.seed(17)
cfunc <- function(n) 
{  
   mySimul <- data.frame()
   for  (i in 1 : nbSimul) 
   { 
     theMean <- mean(rexp(n, lambda))
     mySimul <-  rbind(mySimul,c(sqrt(n)*(theMean - theoricalMean)/ theoricalStandardDeviation,n))
   }
   names(mySimul) <- c('x','size')
   return(mySimul)
}   
dat <- data.frame(  rbind(cfunc(10),cfunc(100)))
                
ggplot(dat, aes(x=x))+
   geom_histogram(aes(y=..density..), colour="black", fill="white")+
   geom_density(alpha=.2, fill="#000066")+facet_grid(. ~ size)
```  
  
We can see a nice curve center on 0 and become more normal with n greater 

## Conclusion : 
1. The theorical mean and variancee are quite near the simulmation mean and variance 
2. With the formula of the CLT we can see a cuvre 
  
$\Rightarrow$ So this distribution is approximately normal.

\newpage  

## Appendix  
Another ways to know if the distribution is approximately normal we can also draw a normal Quantile-Quantile plot or use the wilcoxon test :

### Quantile-Quantile plot
```{r}
qqnorm(mySimul$valueSim, col = 'blue',main = 'Distribution of the simulation')
qqline(mySimul$valueSim, col = 'red')
```  

With the normal Quantile-Quantile plot the blue circles follow the red line more or less at the extrems but follow it on the middle 
$\Rightarrow$ So this distribution is approximately normal.
  
### wilcoxon test
```{r}
wilcox.test(mySimul$valueSim)
```

the p-value of the wilconox test is less than 5% 
$\Rightarrow$ So this distribution is approximately normal.

## End of the document  